Simulation Modeling of Glycosylation Effects on Potassium Channels of Mouse Cardiomyocytes

### 2021-06-21 draft.tex
Renamed

% === ABSTRACT ==========================================================
% =======================================================================
\begin{abstract}
Cardiovascular disease is a leading cause of death worldwide. Dilated cardiomyopathy (DCM) is the third most common cause of heart failure and the most frequent reason for heart transplantation; upward of 70\% of DCM cases are considered idiopathic. We showed, through in-vitro experiments, that reduced hybrid/complex N-glycosylation in mouse cardiomyocytes is sufficient to cause DCM. Further, we investigated the role of reduced complex N-glycosylation in the gating and activities of voltage-gated ion channels, specifically, the voltage-gated $\text{Na}^{+}$ ($\text{Na}_{v}$), $\text{Ca}^{2+}$ ($\text{Ca}_{v}$), and $\text{K}^{+}$ ($\text{K}_{v}$) channels responsible for ventricular myocyte action potentials (APs). In addition to direct effects of reduced N-glycosylation on $\text{Na}_{v}$ and $\text{Ca}_{v}$ gating, we observed aberrant electrical signalings such as prolonged APs and abnormal early re-activations, consistent with an observed reduction in $\text{K}_{v}$ expression and activity. However, it is difficult to rigorously determine the effects of reduced hybrid/complex N-glycosylation on $\text{K}_{v}$ activity, as there are multiple isoforms in the mouse ventricle including, $\text{K}_{v}1.5$, $\text{K}_{v}2.1$, and $\text{K}_{v}4.2$, that collectively produce the repolarizing phase of the AP. Experimentally, because of the unique inactivation kinetics of the $\text{K}_{v}$ isoforms, only the sum of $\text{K}^{+}$ currents ($\text{I}_{Ksum}$) can be recorded and decomposed into component currents using exponential fitting to estimate the currents flowing through $\text{K}_{v}$ isoforms. However, such decomposition cannot describe effectively the mechanism of $\text{K}_{v}$ under different conditions. Here, we propose the use of simulation models to describe statistics and data recorded from in-vitro experiments. We developed simulation models for dominant $\text{K}^{+}$ currents in mouse ventricular myocytes and calibrate their parameters using the in-vitro data under normal and reduced glycosylation conditions through ablation of the Mgat1 gene. A novel model calibration procedure is proposed to deal with the complex structure of the ionic current models based on nonlinear functions and differential equations. Experimental results show that the proposed method has strong potential to build complex simulation models with sparse data and predict characteristics of multiple $\text{K}_{v}$ isoforms to understand detailed kinetics.
\end{abstract}

@@

% === I. INTRODUCTION ===================================================
% =======================================================================
\section{Introduction}
\IEEEPARstart{H}{eart} disease is the leading cause of death globally, accounting for 23\% of deaths in the U.S. in 2017 \cite{cdc2019deaths}. Dilated cardiomyopathy (DCM) is one of the most common cause of heart failure that calls for heart transplantation \cite{weintraub2017dilated}. DCM is characterized by enlarged and weakened ventricular chambers, and it is associated with systolic and contractile dysfunction that has a high risk to heart failure, with approximately 70\% of DCM cases regarded as idiopathic \cite{weintraub2017dilated, lakdawala2013dilated, hershberger2011update}. There has been consistent and increasing evidence of a link between aberrant glycosylation and heart failure \cite{barrans2002global, hwang2002microarray, yung2004gene, nishio1995identification, knezevic2009variability, miura2016glycomics, nagai2016aberrant, yang2015glycoproteins, jacob2013altered, ufret2001role, gehrmann2003cardiomyopathy, marques2017cardiac}. Recently, we showed that reduction of hybrid/complex N-glycosylation in mouse cardiomyocytes through ablation of the Mgat1\footnote{Mannosyl ($\alpha$-1,3-)-Glycoprotein $\beta$-1,2-N-acetylglucosaminyltransferase} gene that encodes a critical glycosyltransferase (GlcNAcT1\footnote{UDP-GlcNAc:$\alpha$-3-D-mannoside-$\beta$1,2-N-acetylglucosaminyltransferase}) (Mgat1KO model), is sufficient to cause DCM \cite{ednie2019reduced, ednie2019reduced2}. Mgat1KO mice develop DCM, heart failure, and 100\% die early, likely from ventricular arrhythmias resulting in sudden cardiac death. Further, Mgat1KO ventricular myocytes demonstrated altered electromechanical functions, including excitation-contraction (EC) coupling and contraction consistencies with observed changes in electrical signaling caused by acute and downstream (disease-related) effects on voltage-gated ion channel (VGIC) gating and activity \cite{ednie2019reduced}.

Thus, we investigated the impact of reduced hybrid/complex N-glycosylation through the lens of electrophysiology. Electrophysiology refers to studies of electrical properties in cells, tissues, or organs, including the examination of voltage, currents, and changes in scales from ion channels to organ systems \cite{scanziani2009electrophysiology}. Electrical signaling in the heart has vital functions related to intra and extracellular communication, rhythmicity of heartbeats, and provides a driving force for contraction. The action potential (AP) is the net transmembrane potential varying over time which results from the composite activities of different ion channels \cite{grant2009cardiac}. Fig. 1-(a) shows the simulated AP of a mouse ventricular apex myocyte using models adopted from \cite{bondarenko2004computer}. The AP is the result of orchestrated activities of various ionic currents described in Fig. 1-(b). Even small changes in VGIC's function can contribute to aberrant AP waveform and/or conduction, leading to arrhythmias. As illustrated in Fig. 2, voltage-gated $\text{Na}^{+}$ channels ($\text{Na}_{v}$) are responsible for AP initiation, or the depolarization phase, voltage-gated $\text{Ca}^{2+}$ channels ($\text{Ca}_{v}$) are responsible for the prolonged depolarization phase particularly in larger species (“plateau”), while several voltage-gated $\text{K}^{+}$ channel ($\text{K}_{v}$) isoforms are collectively responsible for AP deactivation, i.e., repolarization. Like most transmembrane proteins, VGICs are heavily glycosylated membrane proteins \cite{ednie2011modulation}. Studies have proven that glycosylation can affect VGIC function predominantly. For example, we reported that a saturating, electrostatic effect of negatively charged sialic acids, which are typically attached to the terminal of glycan branches, significantly altered cardiac electrical signaling \cite{ednie2013expression, ednie2015reduced}. In our recent studies \cite{ednie2019reduced, ednie2019reduced2}, we observed, for Mgat1KO ventricular apex myocytes, aberrant electrical signaling such as prolonged APs and abnormal early re-activations, caused by direct and disease-related effects of reduced N-glycosylation on VGIC gating and activity that are consistent with the AP waveform and conduction anomalies observed.
\begin{figure}
    \label{fig1}
    \centering
    \subfloat[]{\includegraphics[width=0.8\linewidth]{photo/sim_ap.png}}
    \\
    \subfloat[]{\includegraphics[width=0.8\linewidth]{photo/sim_currents.png}}
    \caption{Simulation results of the (a) action potential and (b) relevant ionic currents (i.e., $\text{I}_{Kto}$, $\text{I}_{Kslow}$, $\text{I}_{Kss}$, $\text{I}_{Na}$, and $\text{I}_{CaL}$) of mouse ventricular myocytes. Potassium currents have positive currents, while sodium and calcium currents are negative.}
\end{figure}

@@

Although these in-vitro experiments showed changes in activities of ion channels, there are certain limitations that include: 1) \textit{Detailed ion channel kinetics are difficult to determine rigorously using in-vitro experiments alone}. From the electrophysiological experiment, pathological electrical signaling can be observed through measurements of APs at the whole-cell level and currents at the ion-channel level. However, it is difficult to relate, rigorously, the impacts of VGIC gating changes to the altered AP waveform/conduction and vice versa. 2) \textit{Segregating different $\text{K}^{+}$ currents ($\text{I}_{K}$) is difficult using whole-cell recording experimental techniques only}. $\text{I}_{Ksum}$ is the result of activity of multiple isoforms with each isoform producing a different, but slightly overlapping (in activation and inactivation voltages) component of the total $\text{K}^{+}$ current ($\text{I}_{Ksum}$) \cite{brouillette2004functional} that thereby contributes to various portions of AP repolarization. However, in-vitro experiments can only measure the sum of these component currents ($\text{I}_{Ksum}$) using whole-cell voltage-clamp protocols. Even pharmacologic separation of $\text{K}_{v}$ current types is difficult, as the specificity of drugs for a single current type is not ideal. Therefore, component $\text{K}^{+}$ currents are usually estimated through multiple exponential fits of the decaying portion (due to channel inactivation) of the total $\text{K}^{+}$ current \cite{brunet2004heterogeneous}. Although the exponential function illustrates the shape of a component-current trace well, it cannot adequately describe kinetic dynamics of the current, nor fully and rigorously distinguish among currents produced by different $\text{K}_{v}$ isoforms because of their slightly overlapping voltage-dependence of gating. As $\text{K}_{v}$ have a critical function in repolarizing the cell, it is imperative to examine $\text{K}_{v}$ isoform thoroughly.

Therefore, in this study, we propose mathematical simulation methods to model different $\text{K}^{+}$ component currents so that $\text{I}_{Ksum}$ is decomposed to study two major voltage-dependent component $\text{K}^{+}$ currents with simulation models, $\text{I}_{Kto}$ and $\text{I}_{Kslow}$. Fig. 3 shows an example of the major component $\text{K}^{+}$ currents with the sum $\text{K}^{+}$ current shown as a red-solid line. Except for the $\text{K}^{+}$ steady-state current ($\text{I}_{Kss}$) component, $\text{K}^{+}$ currents have the shape of reaching the peak relatively early following a depolarization that decreases at variable rates. Thus, there are two critical measures for the shape of $\text{K}^{+}$ current traces: peak value and the current attenuation. As reported in \cite{ednie2019reduced}, we derived key statistics of the component currents, amplitudes, and inactivation rates ($\tau$'s) using bi-exponential fits from the $\text{I}_{Ksum}$ traces of the in-vitro experimental data measured following a 4.5 s stimulus depolarizing pulse. Amplitude and $\tau$ refer to the value of the peak and the time constant when current reduces by $e^{-1}$ (almost 63\%) of the peak. These key statistics were used to calibrate biophysical simulation models. The outcomes and impacts of this research are summarized as follows.
\begin{itemize}
    \item Detailed $\text{K}_{v}$ kinetics can be simulated and predicted using the proposed model to get a better understanding of $\text{K}_{v}$ kinetics, unlike an exponential function, which only describes a current trace.
    \item The simulation models can be fitted to current trace data directly, from which the differences across observations, i.e., cells, can be characterized.
\end{itemize}

% === II. RESEARCH BACKGROUND ===========================================
% =======================================================================
\section{Research Background}
\subsection{Glycosylation and DCM}
Protein glycosylation is an essential cellular process that impacts many cell functions \cite{marques2017cardiac}. Briefly, protein glycosylation is the sequential co-/post-translational process of attaching sugar residues (glycans) to proteins. Cardiac VGICs are heavily glycosylated proteins with upwards of 30\% of their mass consisting of N- and O-linked glycans. A growing number of cardiac diseases, including DCM and hypertrophic cardiomyopathy, can present with concurrent, albeit, modest changes in glycosylation \cite{gehrmann2003cardiomyopathy, footitt2009cardiomyopathy, marques2017cardiac}. Mgat1 expression was implicated in cardiac function and shown to be down-regulated in human end-stage idiopathic DCM. Genome-wide searches identified changes in glycosylation-related gene expression in human idiopathic DCM, including glycosyltransferases \cite{barrans2002global, hwang2002microarray, yung2004gene}; and proteomic/glycomic studies show changes in serum N-glycosylation in heart disease models and in humans with DCM risk factors \cite{nishio1995identification, knezevic2009variability, miura2016glycomics, nagai2016aberrant, yang2015glycoproteins}. Models of DCM/heart failure were associated with subtle changes in glycosylation of proteins involved in electromechanical processes, and ~20\% of patients with congenital disorders of glycosylation (CDG) present with cardiac deficits, including idiopathic DCM \cite{gehrmann2003cardiomyopathy, marques2017cardiac}. The data suggest a correlation between modest changes in extracellularly facing glycosylation and DCM/heart disease \cite{yung2004gene, hwang2002microarray, barrans2002global, nagai2016aberrant, yang2015glycoproteins}.

\subsection{Simulation Modeling of Mouse Ventricular Myocytes}
Electrophysiology was pioneered by Hodgkin and Huxley with a series of famous experiments on the squid giant axon \cite{hodgkin1952quantitative}. Their simulation model was built upon two fundamental principles observed in the experimental data. 1) Cells are selectively activated by different ion channels. 2) Activities of ion channels to produce ionic currents are controlled through voltage changes, from which the name ``voltage-gated ion channels'' originates. The voltage-gating mechanism is embedded in the simulations model by partial differential equations. Not only are simulation models compatible with data and descriptive, but they also simulate changes of action potential waveforms, and cell conductance. Thus, models provide connections among ionic currents, action potentials, and conductance. However, significant domain knowledge in kinetics of ion channels and trial-and-error investigations are required, as advanced cardiomyocyte simulation models have complex structures based on nonlinear functions and differential equations with high dimensionality.

In the literature, optimization procedures have been proposed to fit cardiomyocyte simulation models to electrophysiological experimental data in an autonomous way as much as possible. $\text{Na}_{v}$ activity was coupled with in-vitro data of mouse cardiomyocyte using fractional factorial design and a Gaussian process model \cite{du2015statistical}. Fractional factorial designs were used to find significant control variables from the entire set of tuning parameters to reduce dimensionality. The Gaussian process model, which is computationally cheaper than the original $\text{Na}_{v}$ model, has served as a surrogate model to find optimal parameters. In \cite{du2017}, $\text{K}^{+}$ component currents, $\text{I}_{Kto}$ ($\text{K}_{v4.2}$ activity), and $\text{I}_{Kslow}$ ($\text{K}_{v1.5}/\text{K}_{v2.1}$ activity) simulation models are calibrated with a nonlinear optimization algorithms, i.e., the trust-region method.

In both studies, statistics such as steady-state activation (SSA), steady-state inactivation (SSI), and time constant of inactivation at different voltage were used to calculate goodness-of-fit as an objective function of optimization to minimize discrepancies with in-vitro experiment data. However, these characteristic statistics, which are derived from current traces recorded during in-vitro experiments, are not always available. For example, in our Mgat1KO data, currents were too small at early-activating voltages to reliably calculate the characteristic statistics.

\subsection{$\text{K}_{v}$ Isoforms}
\begin{figure}
    \label{fig3}
    \centering
    \includegraphics[width=0.8\linewidth]{photo/k_decomp_eg.png}
    \caption{Example of $\text{K}^{+}$ current decomposition into three component currents. Each component $\text{K}^{+}$ current has different characteristics in terms of their peak and decaying phase}
\end{figure}
Activation of several different $\text{K}_{v}$ isoforms produces multiple component currents, as illustrated in Fig. 1-(b). However, it is difficult to rigorously delineate the multiple component $\text{K}^{+}$ currents through voltage-clamp experiments alone as the sum of $\text{K}^{+}$ currents is measured \cite{brouillette2004functional}. Thus, $\text{I}_{Ksum}$ is usually decomposed mathematically by fitting the peak and decaying portion of each component current trace with an exponential function \cite{brunet2004heterogeneous}. As given in Equation (1), a standard exponential function has three parameters: amplitude ($\text{A}$), time constant ($\tau$), and constant offset value $\text{C}$. $t$ is time in millisecond. Amplitude refers to the value of the peak and $\tau$ to the time constant when current reduces by $e^{-1}$ (~63\%) of the peak.

@@

As mentioned earlier, $\text{I}_{Ksum}$ was measured using whole-cell voltage clamp protocols. Cells were held at -70 mV and then depolarized by 10 mV voltage steps from -50 mV to 50 mV for 4.5 seconds. Although our experimental data also include 25 s depolarizations, which allows distinguishing $\text{I}_{Kslow1}$ ($\text{K}_{v}1.5$ activity) and $\text{I}_{Kslow2}$ ($\text{K}_{v}2.1$ activity) from $\text{I}_{Kslow}$, these data are not studied in this paper. Observations made at only one voltage step make it difficult to verify models and thus will be studied in future works.

Fig. 4, adopted from \cite{ednie2019reduced}, shows our in-vitro experimental results for 50 mV voltage steps. The peaks of two component currents, $\text{I}_{Kto}$ and $\text{I}_{Kslow}$ are reduced in the Mgat1KO group. Steady-state current $\text{I}_{Kss}$, remaining constant during the depolarization, is also reduced in the Mgat1KO group. To be specific, $\text{A}_{Kslow}$ is most reduced compared to WT, by ~77\%. In addition, $\tau_{Kslow}$ of the Mgat1KO group is ~30\% larger than the WT group, which means $\text{I}_{Kslow}$ is prolonged with reduced N-glycosylation. However, $\tau_{Kto}$ does not show significant difference.
\begin{figure}
    \label{fig4}
    \centering
    \subfloat[]{\includegraphics[width=1\linewidth]{photo/barplot_A_star.png}}
    \\
    \subfloat[]{\includegraphics[width=1\linewidth]{photo/barplot_tau_star.png}}
    \caption{In-vitro experimental results, adopted from \cite{ednie2019reduced}, of averages and error bars of (a) the peaks and constant current, and (b) inactivation time constant. As reported in \cite{ednie2019reduced}, significant differences between WT and Mgat1KO at $p \leq 0.05$ are indicated by an *. ($n=35$ for WT and $n=38$ for Mgat1KO.}
\end{figure}

% === IV. RESEARCH METHODOLOGY ==========================================
% =======================================================================
\section{Research Methodology}
In this study, we propose an optimization procedure to calibrate simulation models of potassium channel isoform activity in mouse cardiomyocytes \cite{ednie2019reduced}. Coupling in-vitro data of $\text{K}_{v}$ with a simulation model has advantages that a bi-exponential function cannot provide. For example, ionic currents, action potentials, and conductance can be determined more rigorously; and models allow for the simulation of conditions different from those used in the in-vitro experiment. In most existing studies, statistics over various voltage steps were derived using the bi-exponential function, i.e., steady-state activation (SSA), steady-state inactivation (SSI), and time constants of inactivation. Then these statistics were used to calibrate simulation models of $\text{I}_{Kto}$ and $\text{I}_{Kslow}$ \cite{du2017}. However, in our electrophysiological experiment, the small current size observed in Mgat1KO myocyte makes it difficult to measure some parameters such as SSA and SSI.

>>>>

% === ABSTRACT ==========================================================
% =======================================================================
\begin{abstract}
Cardiovascular disease is a leading cause of death worldwide. Dilated cardiomyopathy (DCM) is the third most common cause of heart failure and the most frequent reason for heart transplantation; upward of 70\% of DCM cases are considered idiopathic. We showed, through in-vitro experiments, that reduced hybrid/complex N-glycosylation in mouse cardiomyocytes is sufficient to cause DCM. Further, we investigated the role of reduced complex N-glycosylation in the gating and activities of voltage-gated ion channels, specifically, $\text{Na}^{+}$ ($\text{Na}_{v}$), $\text{Ca}^{2+}$ ($\text{Ca}_{v}$), and $\text{K}^{+}$ ($\text{K}_{v}$) channels responsible for ventricular myocyte action potentials (APs). In addition to direct effects of reduced N-glycosylation on $\text{Na}_{v}$ and $\text{Ca}_{v}$ gating, we observed aberrant electrical signalings such as prolonged APs and abnormal early re-activations, consistent with an observed reduction in $\text{K}_{v}$ expression and activity. However, it is difficult to rigorously determine the effects of reduced hybrid/complex N-glycosylation on $\text{K}_{v}$ activity, as there are multiple isoforms in the mouse ventricle including, $\text{K}_{v}1.5$, $\text{K}_{v}2.1$, and $\text{K}_{v}4.2$, that collectively produce the repolarizing phase of the AP. Experimentally, because of the unique inactivation kinetics of $\text{K}_{v}$ isoforms, only the sum of $\text{K}^{+}$ currents ($\text{I}_{Ksum}$) can be recorded and decomposed using exponential fitting to estimate component currents flowing through $\text{K}_{v}$ isoforms. However, such decomposition cannot describe effectively the mechanism of $\text{K}_{v}$ under different glycosylation conditions. Here, we propose to develop simulation models for dominant $\text{K}^{+}$ currents in mouse ventricular myocytes and calibrate their parameters using the in-vitro data under normal and reduced glycosylation conditions through ablation of the Mgat1 gene. Calibrated models help describe key characterisitcs of glycosylation effects and data recorded from in-vitro experiments. A novel model calibration procedure is proposed to deal with complex ionic current models based on nonlinear functions and differential equations. Experimental results show that the proposed method has strong potential to calibrate complex simulation models with sparse data and predict characteristics of multiple $\text{K}_{v}$ isoforms to understand detailed kinetics.
\end{abstract}

@@

% === I. INTRODUCTION ===================================================
% =======================================================================
\section{Introduction}
\IEEEPARstart{H}{eart} disease is the leading cause of death globally, accounting for 23\% of deaths in the U.S. in 2017 \cite{cdc2019deaths}. Dilated cardiomyopathy (DCM) is one common cause of heart failure that calls for heart transplantation \cite{weintraub2017dilated}. DCM is characterized by enlarged and weakened ventricular chambers, and it is associated with systolic and contractile dysfunction that has a high risk to heart failure, with approximately 70\% of DCM cases regarded as idiopathic \cite{weintraub2017dilated, lakdawala2013dilated, hershberger2011update}. There has been consistent and increasing evidence of a link between aberrant glycosylation and heart failure \cite{gehrmann2003cardiomyopathy, footitt2009cardiomyopathy, marques2017cardiac}. Recently, we showed that reduction of hybrid/complex N-glycosylation in mouse cardiomyocytes, through ablation of the Mgat1\footnote{Mannosyl ($\alpha$-1,3-)-Glycoprotein $\beta$-1,2-N-acetylglucosaminyltransferase} gene that encodes a critical glycosyltransferase (GlcNAcT1\footnote{UDP-GlcNAc:$\alpha$-3-D-mannoside-$\beta$1,2-N-acetylglucosaminyltransferase}) (Mgat1KO model), is sufficient to cause DCM \cite{ednie2019reduced, ednie2019reduced2}. Mgat1KO mice develop DCM, heart failure, and 100\% die early, likely from ventricular arrhythmias resulting in sudden cardiac death. Further, Mgat1KO ventricular myocytes demonstrated altered electromechanical functions, including excitation-contraction (EC) coupling, are consistent with observed changes in electrical signaling caused by acute and downstream (disease-related) effects on voltage-gated ion channel (VGIC) gating and activity \cite{ednie2019reduced}.

Thus, we investigated the impact of reduced hybrid/complex N-glycosylation through the lens of electrophysiology. Electrophysiology refers to studies of electrical properties in cells, tissues, or organs, including the examination of voltage, currents, and changes in scales from ion channels to organ systems \cite{scanziani2009electrophysiology}. Electrical signaling in the heart has vital functions related to intra and extracellular communication, rhythmicity of heartbeats, and provides a driving force for contraction. The action potential (AP) is the net transmembrane potential varying over time which results from the composite activities of different ion channels \cite{grant2009cardiac}. Fig. 1-(a) shows the simulated AP of a mouse ventricular apex myocyte using models adopted from \cite{bondarenko2004computer}. The AP is the result of orchestrated activities of various ionic currents described in Fig. 1-(b). Even small changes in VGIC's function can contribute to aberrant AP waveform and/or conduction, leading to arrhythmias. As illustrated in Fig. 2, voltage-gated $\text{Na}^{+}$ channels ($\text{Na}_{v}$) are responsible for AP initiation, or the depolarization phase, voltage-gated $\text{Ca}^{2+}$ channels ($\text{Ca}_{v}$) are responsible for the prolonged depolarization phase particularly in larger species (“plateau”), while several voltage-gated $\text{K}^{+}$ channel ($\text{K}_{v}$) isoforms are collectively responsible for AP deactivation, i.e., repolarization. Like most transmembrane proteins, VGICs are heavily glycosylated membrane proteins \cite{ednie2011modulation}. Studies have proven that glycosylation can affect VGIC function predominantly. For example, we reported that a saturating, electrostatic effect of negatively charged sialic acids, which are typically attached to the terminal of glycan branches, significantly altered cardiac electrical signaling \cite{ednie2013expression, ednie2015reduced}. Our recent studies observed aberrant electrical signaling, e.g., prolonged APs and abnormal early re-activations, in Mgat1KO ventricular apex myocytes \cite{ednie2019reduced, ednie2019reduced2}. These observations strongly support direct and disease-related effects of reduced N-glycosylation on VGIC gating and activity.
\begin{figure}
    \label{fig1}
    \centering
    \subfloat[]{\includegraphics[width=0.8\linewidth]{photo/sim_ap.png}}
    \\
    \subfloat[]{\includegraphics[width=0.8\linewidth]{photo/sim_currents.png}}
    \caption{Simulation results of the (a) action potential and (b) relevant ionic currents (i.e., $\text{I}_{Kto}$, $\text{I}_{Kslow}$, $\text{I}_{Kss}$, $\text{I}_{Na}$, and $\text{I}_{CaL}$) of mouse ventricular myocytes. Potassium currents have positive currents, while sodium and calcium currents are negative.}
\end{figure}

@@

Although these in-vitro experiments showed changes in activities of ion channels, there are certain limitations that include: 1) \textit{Detailed ion channel kinetics are difficult to determine rigorously using in-vitro experiments alone}. From the electrophysiological experiment, pathological electrical signaling can be observed through measurements of APs at the whole-cell level and currents at the ion-channel level. However, it is difficult to relate, rigorously, the impacts of VGIC gating changes to the altered AP waveform/conduction and vice versa. 2) \textit{Segregating different $\text{K}^{+}$ currents ($\text{I}_{K}$) is difficult using whole-cell recording experimental techniques only}. $\text{I}_{Ksum}$ is the result of activity of multiple isoforms with each isoform producing a different, but slightly overlapping (in activation and inactivation voltages) component of the total $\text{K}^{+}$ current ($\text{I}_{Ksum}$) \cite{du2017} that thereby contributes to various portions of AP repolarization. However, in-vitro experiments can only measure the sum of these component currents ($\text{I}_{Ksum}$) using whole-cell voltage-clamp protocols. Even pharmacologic separation of $\text{K}_{v}$ current types is difficult, as the specificity of drugs for a single current type is not ideal. Therefore, component $\text{K}^{+}$ currents are usually estimated through multiple exponential fits of the decaying portion (due to channel inactivation) of the total $\text{K}^{+}$ current \cite{brunet2004heterogeneous}. Although the exponential function illustrates the shape of a component-current trace well, it cannot adequately describe kinetic dynamics of the current, nor fully and rigorously distinguish among currents produced by different $\text{K}_{v}$ isoforms because of their slightly overlapping voltage-dependence of gating. As $\text{K}_{v}$ have a critical function in repolarizing the cell, it is imperative to model and examine $\text{K}_{v}$ isoform thoroughly.

Therefore, this paper presents simulation modeling of dominant $\text{K}^{+}$ currents in mouse ventricular myocytes and further calibrate their parameters using the in-vitro data under normal and reduced glycosylation conditions through ablation of the Mgat1 gene. Calibrated models help describe key characteristics of glycosylation effects and data recorded from in-vitro experiments \cite{ednie2019reduced}. Specifically, we propose an optimization procedure to calibrate simulation models of potassium channel isoform activity, $\text{I}_{Kto}$ and $\text{I}_{Kslow}$, in mouse cardiomyocytes.  Fig. 3 shows  major component $\text{K}^{+}$ currents, including $\text{I}_{Ksum}$, $\text{I}_{Kto}$, $\text{I}_{Kslow}$ and $\text{I}_{Kss}$. Key characteristics of component currents, amplitudes, and inactivation rates ($\tau$'s) are measured from in-vitro experimental data with a 4.5 s voltage-clamp protocol (see details in Section III). The major contributions of this paper are summarized as follows.
\begin{itemize}
    \item Simulation models are integrated with in-vitro experiments to investigate the effects of reduced N-glycosylation on $\text{K}_{v}$ activity of mouse myocytes. Computer simulation helps demonstrate the voltage-gating mechanism and conductance kinetics that cannot be readily available through traditional exponential models. 
    \item A self-breeding Genetic Algorithm procedure is developed to calibrate simulation models of voltage-gated $\text{K}^{+}$ channel of mouse ventricular apex myocytes based on key statistics and raw data obtained from in-vitro experiments.
    \item Calibrated models facilitate the prediction of $\text{K}_{v}$ characteristics at different voltages that are not observed in the in-vitro experiments. In the Mgat1KO group, both $\text{I}_{Kto}$ and $\text{I}_{Kslow}$ densities are shown to be significantly reduced and the rate of $\text{I}_{Kslow}$ inactivation is much slower.
    \item Simulation modeling of cardiac myocytes is conducive to gain a better understanding of detailed $\text{K}_{v}$ kinetics, as well as how reduced glycosylation through  ablation of Mgat1 gene impacts $\text{K}_{v}$ kinetics.
\end{itemize}

% === II. RESEARCH BACKGROUND ===========================================
% =======================================================================
\section{Research Background}
\subsection{Glycosylation and DCM}
Protein glycosylation is an essential cellular process that impacts many cell functions \cite{marques2017cardiac}. Briefly, protein glycosylation is the sequential co-/post-translational process of attaching sugar residues (glycans) to proteins. Cardiac VGICs are heavily glycosylated proteins with upwards of 30\% of their mass consisting of N- and O-linked glycans. A growing number of cardiac diseases, including DCM and hypertrophic cardiomyopathy, can present with concurrent, albeit, modest changes in glycosylation \cite{gehrmann2003cardiomyopathy, footitt2009cardiomyopathy, marques2017cardiac}. Mgat1 expression was implicated in cardiac function and shown to be down-regulated in human end-stage idiopathic DCM. Genome-wide searches identified changes in glycosylation-related gene expression in human idiopathic DCM, including glycosyltransferases \cite{barrans2002global, hwang2002microarray, yung2004gene}; and proteomic/glycomic studies show changes in serum N-glycosylation in heart disease models and in humans with DCM risk factors \cite{nishio1995identification, knezevic2009variability, miura2016glycomics, nagai2016aberrant, yang2015glycoproteins}. Models of DCM/heart failure were associated with subtle changes in glycosylation of proteins involved in electromechanical processes, and ~20\% of patients with congenital disorders of glycosylation (CDG) present with cardiac deficits, including idiopathic DCM \cite{gehrmann2003cardiomyopathy, marques2017cardiac}. The data suggest a correlation between modest changes in extracellularly facing glycosylation and DCM/heart disease.

\subsection{Simulation Modeling of Mouse Ventricular Myocytes}
Electrophysiology was pioneered by Hodgkin and Huxley with a series of well-known experiments on the squid giant axon \cite{hodgkin1952quantitative}. Their simulation model was built upon two fundamental principles observed in the experimental data. 1) Cells are selectively activated by different ion channels. 2) Activities of ion channels to produce ionic currents are controlled through voltage changes, from which the name ``voltage-gated ion channels'' originates. The voltage-gating mechanism is embedded in the simulations model by partial differential equations. Not only are simulation models compatible with data and descriptive, but they also simulate changes of action potential waveforms, and cell conductance. Thus, models provide connections among ionic currents, action potentials, and conductance. However, significant domain knowledge in kinetics of ion channels and trial-and-error investigations are required, as advanced cardiomyocyte simulation models have complex structures based on nonlinear functions and differential equations with a high-dimensional set of parameters.

In the literature, optimization procedures have been proposed to fit cardiomyocyte simulation models to electrophysiological experimental data in an autonomous way as much as possible. $\text{Na}_{v}$ activity was coupled with in-vitro data of mouse cardiomyocyte using fractional factorial design and a Gaussian process model \cite{du2015statistical}. Fractional factorial designs were used to find significant control variables from the entire set of tuning parameters to reduce dimensionality. The Gaussian process model, which is computationally cheaper than the original $\text{Na}_{v}$ model, has served as a surrogate model to find optimal parameters. Characteristic curves such as steady-state activation (SSA), steady-state inactivation (SSI), and time constant of inactivation at different voltage were used to calculate goodness-of-fit as an objective function of optimization to minimize discrepancies with in-vitro experiment data. However, they are derived from current traces recorded during in-vitro experiments and may not always be available. For example, in our Mgat1KO data, currents were too small at early-activating voltages to reliably calculate the characteristic statistics.

\subsection{$\text{K}_{v}$ Isoforms}
\begin{figure}
    \label{fig3}
    \centering
    \includegraphics[width=0.8\linewidth]{photo/k_decomp_eg.png}
    \caption{Example of $\text{K}^{+}$ current decomposition into three component currents. Each component $\text{K}^{+}$ current has different characteristics in terms of their peak and decaying phase}
\end{figure}
The orchestrated activity of different $\text{K}_{v}$ isoforms produces multiple component currents, as illustrated in Fig. 1-(b). However, it is difficult to rigorously delineate the multiple component $\text{K}^{+}$ currents through voltage-clamp experiments alone as the sum of $\text{K}^{+}$ currents is measured \cite{brouillette2004functional}. Thus, $\text{I}_{Ksum}$ is usually decomposed mathematically by fitting the peak and decaying portion of each component current trace with an exponential function \cite{brunet2004heterogeneous}. As given in Equation (1), a standard exponential function has three parameters: amplitude ($\text{A}$), time constant ($\tau$), and constant offset value $\text{C}$. $t$ is time in millisecond. Amplitude refers to the value of the peak and $\tau$ to the time constant when current reduces by $e^{-1}$ (~63\%) of the peak.

@@

As mentioned earlier, $\text{I}_{Ksum}$ was measured using whole-cell voltage clamp protocols. Cells were held at -70 mV and then depolarized by 10 mV voltage steps from -50 mV to 50 mV for 4.5 seconds. Note that this paper does not include the 25s depolarization, which often helps distinguish $\text{I}_{Kslow1}$ ($\text{K}_{v}1.5$ activity) and $\text{I}_{Kslow2}$ ($\text{K}_{v}2.1$ activity). Our ongoing experiments focus on voltage-clamp experiments on 25-second periods. The resulted data will be further studied and calibrated with simulation models in future work.

Fig. 4 shows our in-vitro experimental data about key statistics on $\text{I}_{Kto}$ and $\text{I}_{Kslow}$. Peaks of both currents are reduced in the Mgat1KO group. Steady-state current $\text{I}_{Kss}$, remaining constant during the depolarization, is also reduced in the Mgat1KO group. To be specific, $\text{A}_{Kslow}$ is most reduced compared to WT, by ~77\%. In addition, $\tau_{Kslow}$ of the Mgat1KO group is ~30\% larger than the WT group, which means $\text{I}_{Kslow}$ is prolonged with reduced N-glycosylation. However, $\tau_{Kto}$ does not show significant difference.
\begin{figure}
    \label{fig4}
    \centering
    \subfloat[]{\includegraphics[width=1\linewidth]{photo/barplot_A_star.png}}
    \\
    \subfloat[]{\includegraphics[width=1\linewidth]{photo/barplot_tau_star.png}}
    \caption{In-vitro experimental results, adopted from \cite{ednie2019reduced}, of averages and error bars of (a) the peaks and constant current, and (b) inactivation time constant. As reported in \cite{ednie2019reduced}, significant differences between WT and Mgat1KO at $p \leq 0.05$ are indicated by an *. ($n=35$ for WT and $n=38$ for Mgat1KO.}
\end{figure}

% === IV. RESEARCH METHODOLOGY ==========================================
% =======================================================================
\section{Research Methodology}
In this study, we propose an optimization procedure to calibrate simulation models of potassium channel isoform activity in mouse cardiomyocytes. Coupling in-vitro data of $\text{K}_{v}$ with a simulation model has advantages that a bi-exponential function cannot provide. For example, ionic currents, action potentials, and conductance can be determined more rigorously; and models allow for the simulation of conditions different from those used in the in-vitro experiment.
